
\documentclass[11pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amsmath} % for mathematical features
\usepackage{enumitem} % for customizing lists
\title{Introductory Macroeconomics for Engineers}
\author{Martin A. Valdez}
\date{IE 1\\
Final Exam Preparation}

\begin{document}
\maketitle

\subsection*{Question 1}

\begin{enumerate}
    \item Why do we study macroeconomics? (0.25 pts)
    \item What is the point of using models in Macroeconomics? (0.5 pts)
    \item What is the difference between macroeconomics and microeconomics? (0.25 pts)
\end{enumerate}

\subsection*{Question 2}
\begin{enumerate}
    \item What is the definition of GDP? (1 pt)
    \item What are the three approaches to measuring GDP? Provide the formula for each approach. (1.5 pt - 0.5 per approach)
    \item What is the difference between nominal and real GDP? (0.5 pt)
\end{enumerate}

\subsection*{Question 3}
Mention at least four of Kaldor's stylized facts. (0.5 pt each)

\subsection*{Question 4}
Suppose that total output (GDP) in an economy is produced using the following production function:
\begin{align*}
    Y = A F(K,N) = A K^{\alpha} L^{1-\alpha},
\end{align*}
Where \(Y\) is total output, \(K\) is capital, \(L\) is labor, \(A\) is total factor productivity, and \(\alpha \in (0,1)\) 
is the capital share of output.

\begin{enumerate}
    \item Show that F has constant returns to scale, i.e., that \(F(\lambda K, \lambda L) = \lambda F(K,L)\) for all \(\lambda > 0\). (1 pt)
    \item Compute \(\frac{\partial Y }{\partial K} \) and \(\frac{\partial Y }{\partial L} \). (1 pt each)
    \item Use the log transformation to compute the growth rate of output, \(g_Y\), in terms of the growth rates of capital, \(g_K\), labor, 
    \(g_L\), and total factor productivity, \(g_A\). (2 pt)
    \item Suppose that output from \(t\) to \(t+1\) grows 10\%. The capital stock grows from 50 to 60 euros, and the labor
    force grows from 10 to 12 workers. Suppose that \(\alpha = 0.3\).
    \begin{enumerate}
        \item Calculate the growth rate of capital and labor. (1 pt each)
        \item Calculate the growth rate of total factor productivity. (1 pt)
    \end{enumerate}
\end{enumerate}

\subsection*{Question 5}
Recall the Solow model, whose dynamics are given by the following equation:
\begin{align*}
    K_{t+1} = sA K_t^{\alpha}L_t^{1-\alpha} + (1-\delta)K_t
\end{align*}
Where \(K_t\) is the capital stock, \(L_t\) is the labor force, \(s\) is the savings rate, \(\delta\) is the depreciation rate, and \(A\) 
is total factor productivity.

\begin{enumerate}
    \item Transform the equation above into a per-worker basis. (2 pts)
    \item Graph the per-worker capital stock, \(k_{t+1}\), as a function of the per-worker capital stock, \(k_t\). (2 pt)
    \item Show the dynamics of the per-worker capital stock, \(k_{t+1}\) in the graph using the curve \(k_{t+1} = k_t\). (1 pt)
    \item Solve for the equilibrium level of capital per worker, \(k^*\). (2 pt)
    \item What is the growth rate of the per-worker capital stock in the steady state? (1 pt)
\end{enumerate}

\subsection{ Bonus Question}
Where does the savings rate parameter, \(s\), come from in the Solow model? (1 pt)

\end{document}